Global optimization method with dual Lipschitz constant estimates for problems with non-convex constraints\thanks{This research was supported by the Russian Science Foundation, project No. 16-11-10150.}

### 2020-05-02 2019/NUMTA_Barkalov/SoftComputing/article.tex
fix some comments of the Reviewer 1

	The methods for solving the Lipschitzian optimization problems constitute a very important area in the development of the global optimization methods and are the objects of investigations of many researchers. The first algorithms of this class were proposed as early as in the 1970s (\cite{Evtushenko1971,Piyavskii1972,Shubert1972,Strongin1970}); since that time, this area has been continuing to develop extensively (see, for example, \cite{Evtushenko2009,Evtushenko2013,Strongin2000,Sergeyev2013,Jones2009}). Note that these global optimization algorithms are, as a rule, oriented towards solving either  unconstrained optimization problems \break (\cite{Paulavicius2014,Sergeyev2015,Pinter1996,Jones1993,Gablonsky2001}) or problems with constraints of a simple structure (\cite{Vaz2009,Paulavicius2016}). The solving of the problems with complex non-convex constraints is usually performed with the use of the penalty functions method (\cite{Stripinis2019,Pillo2012,Pillo2016}), which has a number of drawbacks.  In particular, this method cannot be applied to the problems with the partially defined functions (i.e. when any constraint is violated, the values of all the other functions of the problem remain undefined).

@@

		The value of the unknown Lipschitz constant estimate affects the convergence rate of the algorithm essentially. Therefore, the issue of its correct estimate is very important. An underestimate of the true value of the constant may lead to the loss of the algorithm convergence to the global solution. At the same time, too high a value of the constant estimate not corresponding to real behavior of the function results in slow convergence of the algorithm to the global minimum point.

		Several typical methods for adaptive estimation of the Lipschitz constant according to the results of the performed search trials are known:
\begin{itemize}
\item global estimate of the constant $L$ in the whole search domain $D$ (\cite{Strongin2000,Pinter1996,Horst1996}).
\item local estimates of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ (\cite{Kvasov2003,Sergeyev2010,Sergeyev2016}).
\item choice of the estimates of the constant $L$ from some set of possible values (\cite{Jones2009,Jones1993,Gablonsky2001,Sergeyev2006}).
\end{itemize}

	Each of the approaches listed above has its own advantages and drawbacks. For example, the use of the global estimate only in the entire search domain may slow down the convergence of the algorithm to the global minimum point. On the other hand, the use of the local estimates accelerating the convergence requires an adequate adjustment of the algorithm parameters in order to ensure the global convergence.

@@

	It is supposed that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_{j}(y)$, $1 \leq j \leq m$, of the constraints satisfy the Lipschitz condition
\begin{align}\label{lipschitz_condition}
	\left|g_j(y')-g_j (y'')\right| \leq L_j \|y'-y''\|, \nonumber \\ 
	y', y''\in D, \; 1\leq j \leq m+1,
\end{align}
with a priori unknown constants $L_j, 1\leq j \leq m+1$, and may be multiextremal.

@@

	Along with the exact solution $y^\ast$ of the problem (\ref{problem}), we will consider also the $\varepsilon$-reserved solution $y_{\epsilon}$, for which 
\small
\begin{align}\label{epsilon_reserved_solution}
	\varphi(y_{\epsilon}) = \min{\left\{\varphi(y):y \in D, g_j(y)\leq -{\epsilon}_j, 1 \leq j \leq m\right\}},
\end{align}
\normalsize
and the set of solutions $Y_{\epsilon}$, which are not worse than the $\varepsilon$-reserved solution in the objective function value, i.e.
\small
\begin{align}\label{Y_epsilon}
	Y_{\epsilon} = \left\{ y \in D, \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_{\epsilon}) \right\}.
\end{align}
\normalsize
Fig. \ref{fig:eps_reserved_solution} illustrates the concepts introduced.

@@

	The existence of the $\varepsilon$-reserved solutions of problem (\ref{problem}) can be interpreted as some analog of the regularity conditions in classical nonlinear programming problems. The applied role of this condition should be also noted. Even if the exact solution $y^\ast$ is found, its practical implementation is possible as some approximation of $y^\ast$ only. Therefore, it is important that there are feasible points from $Q_{m+1}$ close enough to $y^\ast$ (in coordinates and values). The existence of the $\varepsilon$-reserved solution guarantees the existence of such points. In this case, the domain $Y_{\epsilon}$ may play the role of a set of suitable approximations.

@@

	According to (\ref{D_sets}) and (\ref{one_dimensional_problem}), the reduced one-dimensional functions $g_j (y(x)), 1 \leq j \leq m+1,$ will also be defined and computable in the corresponding intervals 
\begin{align}\label{Q_intervals}
	Q_1=[0,1], \; Q_{j+1}=\left\{x \in Q_j : g_j(y(x)) \leq 0 \right\}, \nonumber \\
	1 \leq j \leq m.
\end{align}
These relations allow introducing a classification of the points $x \in [0,1]$ according to the number $\nu = \nu(x)$ of the constraints computable at the given point. We will call this number $\nu = \nu(x)$ the \textit{index} of the point $x$. The index can be determined from the conditions 
$$
	g_j(y(x)) \leq 0, \; 1 \leq j \leq \nu-1, \; g_{\nu}(y(x))>0,
$$
where the last inequality is insufficient if $\nu = m+1$.

@@

	An approach, in which two estimates of the Lipschitz constant for each function are used in the rules of the algorithm, is promising. One of these estimates is significantly greater than the other, which corresponds to the use of two sets of parameters $r_{\nu}^{glob}$ and $r_{\nu}^{loc}$ in the algorithm, where $r_{\nu}^{glob}>r_{\nu}^{loc}>1, 1 \leq \nu \leq m+1$. When using the parameter $r_{\nu}^{loc}$ we shall deal with the smaller estimate of the Lipschitz constant, and using the parameter $r_{\nu}^{glob}$ will correspond to the larger one.

@@

	In Fig. \ref{fig:example_strongin}, the search domain (square) is shown, the feasible domain is highlighted by green, and the trial points generated by IA (a) and IA-DL (b) are marked. The best point is marked by red. For both methods, the evolvent with the parameter $m=10$ was used, the accuracy of the search for the solution $\epsilon=0.002$; the adaptive reserve parameter $\delta=0.008$ from (\ref{epsilon_nu}). For the IA method, the parameters $r_{\nu}=2.3, 1 \leq \nu \leq 4$, were used. The same values were used for the IA-DL method as the parameters $r_{\nu}^{glob}, 1 \leq \nu \leq 4$. Also, the parameters $r_{\nu}^{loc}=1.5, 1 \leq \nu \leq 4,$ were used. To solve the problem with a predefined accuracy $\epsilon=0.002$, 478 trials were required for the IA method compared to only 303 trials for the IA-DL method. The results of larger-scale numerical experiments are presented in Section \ref{sec:4}.

>>>>

	The methods for solving the Lipschitzian optimization problems constitute a very important area in the development of the global optimization methods and are the objects of investigations of many researchers. The first algorithms of this class were proposed as early as in the 1970s (\cite{Evtushenko1971,Piyavskii1972,Shubert1972,Strongin1970}); since that time, this area has been continuing to develop extensively (see, for example, \cite{Evtushenko2009,Evtushenko2013,Strongin2000,Sergeyev2013,Jones2009}). Note that these global optimization algorithms are, as a rule, oriented towards solving either  unconstrained optimization problems \break (\cite{Paulavicius2014,Sergeyev2015,Pinter1996,Jones1993,Gablonsky2001}) or problems with constraints of a simple structure (\cite{Vaz2009,Paulavicius2016}). The solving of the problems with complex non-convex constraints is usually performed with the use of the penalty functions method (\cite{Stripinis2019,Pillo2012,Pillo2016}), which has a number of drawbacks.  In particular, this method cannot be applied to the problems with the partially defined functions (i.e. some of the functions of the problem are defined only at those points which satisfy a selection of constraints).

@@

		The value of the unknown Lipschitz constant estimate affects the convergence rate of the algorithm essentially. Therefore, the issue of its correct estimate is very important. An underestimate of the true value of the constant may lead to the loss of the algorithm convergence to the global solution. At the same time, a too high value of the constant estimate not corresponding to real behavior of the function results in slow convergence of the algorithm to the global minimum point.

		Several typical methods for adaptive estimation of the Lipschitz constant according to the results of the performed search trials are known:
\begin{itemize}
\item global estimate of the constant $L$ in the whole search domain $D$ (\cite{Strongin2000,Pinter1996,Horst1996}).
\item local estimates of the constants $L_i$ in different subdomains $D_i$ of the search domain $D$ (\cite{Sergeyev1998,Kvasov2003,Sergeyev2010,Sergeyev2016}).
\item choice of the estimates of the constant $L$ from some set of possible values (\cite{Jones2009,Jones1993,Gablonsky2001,Sergeyev2006}).
\end{itemize}

	Each of the approaches listed above has its own advantages and drawbacks. For example, the only use of the global estimate in the entire search domain may slow down the convergence of the algorithm to the global minimum point. On the other hand, the use of the local estimates accelerating the convergence requires an adequate adjustment of the algorithm parameters in order to ensure the global convergence.

@@

	It is supposed that the objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and the left-hand sides $g_{j}(y)$, $1 \leq j \leq m$, of the constraints satisfy the Lipschitz condition
\begin{align}\label{lipschitz_condition}
	\left|g_j(y')-g_j (y'')\right| \leq L_j \|y'-y''\|, \nonumber \\ 
	y', y''\in D, \; 1\leq j \leq m+1,
\end{align}
with a priori unknown constants $L_j, \; 0<L_j<\infty, \; 1\leq j \leq m+1$, and may be multiextremal.

@@

	Along with the exact solution $y^\ast$ of the problem (\ref{problem}), we will consider also the $\varepsilon$-reserved solution $y_{\epsilon}$, for which 
\small
\begin{align}\label{epsilon_reserved_solution}
	\varphi(y_{\epsilon}) = \min{\left\{\varphi(y):y \in D, g_j(y)\leq -{\epsilon}_j, 1 \leq j \leq m\right\}},
\end{align}
\normalsize
where the algorithm parameters $\epsilon_j, \; 1\leq j \leq m,$ are non negative numbers, and the set of solutions $Y_{\epsilon}$, which are not worse than the $\varepsilon$-reserved solution in the objective function value, i.e.
\small
\begin{align}\label{Y_epsilon}
	Y_{\epsilon} = \left\{ y \in D, \; g_j(y) \leq 0, \; 1 \leq j \leq m, \; \varphi(y) \leq \varphi(y_{\epsilon}) \right\}.
\end{align}
\normalsize
Fig. \ref{fig:eps_reserved_solution} illustrates the concepts introduced.

@@

	The existence of the $\varepsilon$-reserved solutions of problem (\ref{problem}) can be interpreted as some analogous of the regularity conditions in classical nonlinear programming problems. The applied role of this condition should be also noted. Even if the exact solution $y^\ast$ is found, its practical implementation is possible as some approximation of $y^\ast$ only. Therefore, it is important that there are feasible points from $Q_{m+1}$ close enough to $y^\ast$ (in coordinates and values). The existence of the $\varepsilon$-reserved solution guarantees the existence of such points. In this case, the domain $Y_{\epsilon}$ may play the role of a set of suitable approximations.

@@

	According to (\ref{D_sets}) and (\ref{one_dimensional_problem}), the reduced one-dimensional functions $g_j (y(x)), 1 \leq j \leq m+1,$ will also be defined and computable in the corresponding intervals 
\begin{align}\label{Q_intervals}
	Q_1=[0,1], \; Q_{j+1}=\left\{x \in Q_j : g_j(y(x)) \leq 0 \right\}, \nonumber \\
	1 \leq j \leq m.
\end{align}
These relations allow introducing a classification of the points $x \in [0,1]$ according to the number $\nu = \nu(x)$ of the constraints computable at the given point. We will call this number $\nu = \nu(x)$ the \textit{index} of the point $x$. The index can be determined from the conditions 
$$
	g_j(y(x)) \leq 0, \; 1 \leq j \leq \nu-1, \; g_{\nu}(y(x))>0,
$$
where the last inequality is not necessary if $\nu = m+1$.

@@

	An approach, in which two estimates of the Lipschitz constant for each function are used in the rules of the algorithm, is promising. One of these estimates is significantly greater than the other, which corresponds to the use of two sets of parameters $r_{\nu}^{glob}$ and $r_{\nu}^{loc}$ in the algorithm, where $r_{\nu}^{glob}>r_{\nu}^{loc}>1$ and $1 \leq \nu \leq m+1$. When using the parameter $r_{\nu}^{loc}$ we shall deal with the smaller estimate of the Lipschitz constant, and using the parameter $r_{\nu}^{glob}$ will correspond to the larger one.

@@

	In Fig. \ref{fig:example_strongin}, the search domain (square) is shown, the feasible domain is highlighted by green, and the trial points generated by IA (a) and IA-DL (b) are marked. The best point is marked by red. For both methods, the evolvent with the parameter $m=10$ was used, the accuracy of the search for the solution was $\epsilon=0.002$; the adaptive reserve parameter $\delta=0.008$ is from (\ref{epsilon_nu}). For the IA method, the parameters $r_{\nu}=2.3, 1 \leq \nu \leq 4$, were used. The same values were used for the IA-DL method as the parameters $r_{\nu}^{glob}, 1 \leq \nu \leq 4$. Also, the parameters $r_{\nu}^{loc}=1.5, 1 \leq \nu \leq 4,$ were used. To solve the problem with a predefined accuracy $\epsilon=0.002$, 478 trials were required for the IA method compared to only 303 trials for the IA-DL method. The results of larger-scale numerical experiments are presented in Section \ref{sec:4}.
